---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# openeddy

This package provides utilities for eddy covariance data handling, 
quality checking (similar to Mauder et al., 2013), processing, summarising 
and plotting. It aims to standardise the automated quality checking and make 
data processing reproducible.

## Installation

1. Install devtools package if not available yet.
```r
install.packages("devtools")
```
1. Install openeddy
```r
devtools::install_github("lsigut/openeddy")
```

## Extended Example

An extended example describing the intended eddy covariance data processing
workflow is available at:

https://github.com/lsigut/EC_workflow

## Short Example

Demonstration of generally applicable `openeddy` functions.

```{r}
library(openeddy)
library(REddyProc)
library(ggplot2)
data(Example_DETha98)
str(Example_DETha98[1:4])
DETha98 <- fConvertTimeToPosix(Example_DETha98, 'YDH', 
                               Year = 'Year', Day = 'DoY', Hour = 'Hour')
rename <- names(DETha98) %in% c("DateTime", "Rg")
names(DETha98)[rename] <- c("timestamp", "GR")
DETha98$qc_NEE <- ifelse(is.na(DETha98$NEE), NA, 0)
DETha98$qc_NEE_lowcov <- 
  apply_thr(DETha98$NEE, c(-0.01, 0.01), "qc_NEE_lowcov", "between")
table(DETha98$qc_NEE_lowcov)
DETha98$qc_NEE_runs <- flag_runs(DETha98$NEE, "qc_NEE_runs")
table(DETha98$qc_NEE_runs)
DETha98$qc_NEE_prelim <- 
  combn_QC(DETha98, 
           c("qc_NEE", "qc_NEE_lowcov", "qc_NEE_runs"), 
           "qc_NEE_prelim", additive = FALSE, na.as = NA)
DETha98$qc_NEE_despikeLF <- 
  despikeLF(DETha98, "NEE", "qc_NEE_prelim", "qc_NEE_despikeLF", 
            light = NULL)
table(DETha98$qc_NEE_despikeLF)

summary_QC(DETha98, 
           c("qc_NEE", "qc_NEE_lowcov", "qc_NEE_runs", "qc_NEE_despikeLF"),
           na.as = c(NA, NA, NA, 0))
summary_QC(DETha98, 
           c("qc_NEE", "qc_NEE_lowcov", "qc_NEE_runs", "qc_NEE_despikeLF"),
           na.as = c(NA, NA, NA, 0), cumul = TRUE, plot = TRUE, flux = "NEE")
DETha98$qc_NEE_composite <- 
  combn_QC(DETha98, 
           c("qc_NEE", "qc_NEE_lowcov", "qc_NEE_runs", "qc_NEE_despikeLF"), 
           "qc_NEE_composite", additive = FALSE, na.as = c(NA, NA, NA, 0))

DETha98[, c("P", "PAR", "Rn")] <- NA
(varnames <- varnames(DETha98))
(units <- openeddy::units(DETha98))
sub <- DETha98$DoY >= 29 & DETha98$DoY < 43
DETha98_sub <- DETha98[sub, ]
openeddy::units(DETha98) <- units
plot_eddy(DETha98_sub, "NEE", "qc_NEE", "qc_NEE_composite", skip = "monthly",
          light = "GR")

despikeLF_plots <- 
  despikeLF(DETha98, "NEE", "qc_NEE_prelim", "qc_NEE_despikeLF", 
            light = NULL, plot = TRUE)$plots
despikeLF_plots$`iter 1`$all$`1998-01-27 - 1998-02-08`
```

## References

Publication describing openeddy is not yet available. When describing 
the proposed quality control scheme, please refer to it as similar to:

Mauder, M., Cuntz, M., Drüe, C., Graf, A., Rebmann, C., Schmid, H.P., 
Schmidt, M., Steinbrecher, R., 2013. A strategy for quality and uncertainty 
assessment of long-term eddy-covariance measurements. Agric. For. Meteorol. 
169, 122-135, https://doi.org/10.1016/j.agrformet.2012.09.006